### 2024-04-08 README.qmd
Update with rmd check status
Add analyis

---
format: gfm
---

@@

```{r}
#| include: false
library(tidyverse)
remotes::install_github("ITSLeeds/opsnap")
```

@@

@@

```{r}
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
d = opsnap:::download_and_read(u)
names(d)
# Old names:
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"
```

```{r}
# Function to clean up column names
d_sample = d[1:50, ]
d_sf = opsnap:::op_geocode(d_sample)
mapview::mapview(d_sf)
```

```{r}
table(d$OFFENCE) |>
  sort()
```

```{r}
d |>
  # Reduce nchar of OFFENCE
  mutate(OFFENCE = stringr::str_sub(OFFENCE, 1, 60)) |>
  group_by(OFFENCE) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d) / 50)|>
  ggplot() +
  geom_bar(aes(OFFENCE)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

>>>>

---
format: gfm
execute: 
  message: false
  warning: false
  echo: false
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ITSLeeds/opsnap/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

@@

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
remotes::install_github("ITSLeeds/opsnap")
```

@@

Data for the following years are provided:

```{r}
#| echo: false
library(rvest)
query = ".file-link a"
url = "https://www.westyorkshire.police.uk/SaferRoadsSubmissions"
files_available = read_html(url) |>
  html_nodes(query) |>
  html_text()
links_available = read_html(url) |>
  html_nodes(query) |>
  html_attr("href")
urls = paste0("https://www.westyorkshire.police.uk", links_available)
file_names = basename(links_available)
tibble::tibble(file_names) |>
  knitr::kable()

# In preparation for reading-in all data:
date_str = format(Sys.Date(), "%Y-%m")
file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
```

```{r}
#| eval: false
#| echo: false
# Download them to raw_data/west-yorkshire:
dir.create("raw_data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
pbapply::pblapply(urls, function(u) {
  download.file(u, paste0("raw_data/west-yorkshire/", basename(u)))
})

# Test with 2nd url:
u = urls[2]
d2 = opsnap:::download_and_read(u)
d_all = purrr::map_df(urls, opsnap:::download_and_read)
dir.create("data/west-yorkshire", recursive = TRUE, showWarnings = FALSE)
# Date to nearest month:
write_csv(d_all, file_name)
```

@@

```{r}
#| eval: false
#| echo: true
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
d = opsnap:::download_and_read(u)
names(d)
# Old names:
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"
# New names:
# [1] "mode"     "make"     "model"    "colour"   "offence"  "district" "disposal"
# [8] "date"     "location"
```

<!-- The data looks like this (first 3 rows shown): -->

```{r}
#| echo: false
d_all = read_csv(file_name)
d_all |>
  head(3) |>
  knitr::kable()
```

There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown below):

```{r}
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "all")
d = d_all |>
  opsnap:::filter_nas()
d_complete_monthly = d |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete")
d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
d_monthly |>
  ggplot() +
  geom_line(aes(month, n, colour = records)) +
  geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
       x = "Date",
       y = "Number of records per month")
```

As shown in the graph above, some records `r round(nrow(d) / nrow(d_all) * 100, 1)`% have missing values for locations (these records did not classify as an offence). Filtering these out leaves `r nrow(d)` complete records.

There are `r unique(d$location) |> length()` unique locations in the data, with the most common locations shown below:

```{r}
#| echo: false
d |>
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

# Geocoding

We provide a function to geocode the records:

```{r}
#| eval: false
#| echo: true
d_sample = d[1:5, ]
d_sf = opsnap:::op_geocode(d_sample)
mapview::mapview(d_sf)
```

```{r}
#| eval: false
#| echo: false
# Geocode all data:
d_geocoded = opsnap:::op_geocode(d)
d_geometries = d_geocoded |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
d_sf = sf::st_sf(
  d,
  geometry = d_geometries$geometry
)
west_yorkshire = pct::get_pct_zones("west-yorkshire")
sf::sf_use_s2(FALSE)
d_sf_wy = d_sf[west_yorkshire, ]
# Sanity check
d_sf_wy |>
  sample_n(1000) |>
  # mapview::mapview()
  plot()
# Save to file:
sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
```

# Analysis

Due to inaccuracy in the geocoding, we only know the locations of the records to within around 500m of each crash (although we can link to specific roads).
We'll present the geographic distribution of crashes using a 500m grid:

```{r}
d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
d_projected = sf::st_transform(d_sf, 27700)
library(terra)
raster_template = rast(ext(d_projected), res = 500)
raster_count = rasterize(d_projected, raster_template, fun = "length")
plot(raster_count)
```

The results show there is one outlier with a very high number of crashes. We can remove this and plot the data again:

```{r}
raster_count_no_outlier = raster_count |>
  clamp(-Inf, 100)
plot(raster_count_no_outlier)
```

<!-- You can query the data downloaded with `opsnap` functions, e.g. as follows (results not shown): -->

```{r}
#| eval: false
#| echo: false
table(d$offence) |>
  sort()
```

<!-- Let's make a plot of the data: -->

```{r}
#| eval: false
#| echo: false
d |>
  # Reduce nchar of offence
  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
  group_by(offence) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d) / 50)|>
  ggplot() +
  geom_bar(aes(offence)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
